%%
% This file is part of the User's Guide to RSS
% It contains the appendix for inputfile specifications
%%

\section{Input-file Specifications}

There are exactly four kinds of input files for the \RSS. This includes the project specification files and also the \Lua-script-files that define the robot behaviour.
\begin{enumerate}
	\item The main projectfile containing information about the model. The extension of this type of file is "`.swarm"'.
	\item A file containing robot information. The extension of this file is "`.robot"'.
	\item A file containing obstacle information. The extension of this file is "`.obstacle"'.
	\item \Lua\ file that describes the robot behavouir. The extension of this file is "`.lua"'.
\end{enumerate}


\subsection{Main projectfile}
The following specifications hold only for the main projectfile (with extension \texttt{.swarm}):
\begin{itemize}
	\item A comment begins with a '\#'.
	\item A line is a comment line (beginning with a '\#'), an empty line or a line containing a variable followed by an equal sign followed by a \emph{quoted} value of this variable. Example:
	\begin{verbatim}
		VAR_1="value"
		VAR_2 = "value"
		VAR_3= "value"
		VAR_4 ="value"
	\end{verbatim}
	\item a variable name has to be of the following form: \texttt{[A-Z0-9\_]$^+$}
\end{itemize}


\subsubsection{Variables}
The main project file contains the variables defined in Tables \ref{tab:mainvars} and \ref{tab:mainvars2}.
	
Also the following should be considered:
\begin{itemize}
	\item The order of the variables in the main project file is not important.
	\item If a variable does not appear in the main projectfile, then its default value will be used if such a default value does exist (otherwise an exception will be thrown while loading the main project-file).
\end{itemize}

\clearpage
\begin{sidewaystable}
\scriptsize
	\begin{tabular}{|l|p{0.3\textwidth}|p{0.3\textwidth}|p{0.2\textwidth}|}
		\hline
		\textbf{Variable name} & \textbf{Possible Values} & \textbf{Description} & \textbf{Default}\\\hline\hline
		\texttt{PROJECT\_NAME} & String & Name of the project & -- \\\hline
% 		\texttt{BATTLEBOX\_SIZE} & width, for instance 100 denotes a box of size $100\times 100\times 100$ & Size of bounding box of initial robot positions\\\hline
		\texttt{COMPASS\_MODEL} & Still needs to be specified by the ASG-Team. For instance \texttt{NO\_COMPASS} & Compass model & FULL\_COMPASS\\\hline
		\texttt{ROBOT\_FILENAME} & For instance \texttt{robot\_file}. The extension of the file must not be appended in this variable. & Filename of the robotfile & same as project file\\\hline
		\texttt{OBSTACLE\_FILENAME} & For instance \texttt{obstacle\_file}.  The extension of the file must not be appended in this variable. & Filename of the robotfile & same as project file\\\hline
		\texttt{STATISTICS\_SUBSETS} & A concatenation of none or more of the following strings: \{ALL\}, \{ACTALL\}, \{INACTALL\}, \{MASTERS\}, \{ACTMASTERS\}, \{INACTMASTERS\},  \{SLAVES\}, \{ACTSLAVES\}, \{INACTSLAVES\} &  Defines the subsets of all robots for which to calculate individual statistical data. E.\,g. ``\{ALL\} \{MASTERS\}'' will produce statistical information on \textit{all} robots as well as on \textit{masters only} & NONE\\\hline
		\texttt{STATISTICS\_TEMPLATE} & One of the following: ``ALL'', ``BASIC'' or ``NONE'' & Identifies the set of informations to calculate for each subset. & ALL\\\hline
		\texttt{STATISTICS\_DATADUMP} & Either ``FULL'' or ``NONE'' & Whether or not detailled information (E.\,g. all robots positions at each event) should be streamed to a file during simulation. & NONE\\\hline
		\texttt{ASG} & \texttt{SYNCHRONOUS}, \texttt{ASYNCHRONOUS} or \texttt{SEMISYNCHRONOUS} & Type of ASG & \texttt{SYNCHRONOUS}\\\hline
		  \texttt{ASYNC\_ASG\_SEED} & unsigned int & Seed for asynchronous ASG, only set if ASG=ASYNCHRONOUS & - \\\hline
		    \texttt{ASYNC\_ASG\_PART\_P} & double & Participation Probability for asynch ASG, only set if ASG = ASYNCHRONOUS & - \\\hline
		 \texttt{ASYNC\_ASG\_TIME\_P} & double & parameter governing the timing of asynch ASG, only set if ASG = ASYNCHRNOUS. The lower this is the more often events happen. & - \\\hline
		 
		\texttt{ROBOT\_CONTROL} &  see section \ref{sec:robotControl} & RobotControl to use & -\\\hline
		\texttt{CAMERA\_POSITION} &  \texttt{x,y,z}, where $x,y,z\in\mathbb{R}$& Initial camera position & \texttt{0,0,0}\\\hline
		\texttt{CAMERA\_DIRECTION} &  \texttt{x,y,z}, where $x,y,z\in\mathbb{R}$& Initial camera direction & \texttt{1,0,0}\\\hline


		 
	\end{tabular}
	\caption{Variables in the main project file}\label{tab:mainvars}
\end{sidewaystable}
\thispagestyle{empty}
\clearpage

\clearpage
\begin{sidewaystable}
\scriptsize
	\begin{tabular}{|l|p{0.3\textwidth}|p{0.3\textwidth}|p{0.1\textwidth}|}
		\hline
		\textbf{Variable name} & \textbf{Possible Values} & \textbf{Description} & \textbf{Default}\\\hline\hline

		 \texttt{MARKER\_REQUEST\_HANDLER\_TYPE} &  element from $\{$\texttt{STANDARD,NONE}$\}$ & Type of Marker Request Handler to use & $\{$\texttt{NONE}$\}$\\\hline
		 
		\texttt{TYPE\_CHANGE\_REQUEST\_HANDLER\_TYPE} &  element from $\{$\texttt{STANDARD,NONE}$\}$ & Type of Type Change Request Handler to use. & $\{$\texttt{NONE}$\}$\\\hline
		
		\texttt{POSITION\_REQUEST\_HANDLER\_TYPE} &  element from $\{$\texttt{VECTOR,NONE}$\}$ & Type of Position Request Handler to use & $\{$\texttt{NONE}$\}$\\\hline

		\texttt{VELOCITY\_REQUEST\_HANDLER\_TYPE} &  element from $\{$\texttt{VECTOR,NONE}$\}$ & Type of Velocity Request Handler to use & $\{$\texttt{NONE}$\}$\\\hline

		\texttt{ACCELERATION\_REQUEST\_HANDLER\_TYPE} &  element from $\{$\texttt{VECTOR,NONE}$\}$ & Type of Acceleration Request Handler to use & v\\\hline
		
		 \texttt{STANDARD\_MARKER\_REQUEST\_HANDLER\_SEED} &  integer & Seed for Marker Request Handler to use & $\{$\texttt{NONE}$\}$\\\hline
		 
		\texttt{STANDARD\_TYPE\_CHANGE\_REQUEST\_HANDLER\_SEED} &   integer & Seed for Type Change Request Handler to use. & -\\\hline
		
		\texttt{POSITION\_REQUEST\_HANDLER\_SEED} &   integer & Seed for Position Request Handler to use & -\\\hline

		\texttt{VELOCITY\_REQUEST\_HANDLER\_SEED} &   integer & Seed for Velocity Request Handler to use & -\\\hline

		\texttt{ACCELERATION\_REQUEST\_HANDLER\_SEED} &   integer & Seed for Acceleration Request Handler to use & -\\\hline
		
		
		
		 \texttt{STANDARD\_MARKER\_REQUEST\_HANDLER\_DISCARD\_PROB} &  element from interval $[0,1]$ & Discard probability for Marker Request Handler to use & -\\\hline
		 
		\texttt{STANDARD\_TYPE\_CHANGE\_REQUEST\_HANDLER\_DISCARD\_PROB} & element from interval $[0,1]$ & Discard probability  for Type Change Request Handler to use. & -\\\hline
		
		\texttt{POSITION\_REQUEST\_HANDLER\_DISCARD\_PROB} & element from interval $[0,1]$ & Discard probability  for Position Request Handler to use & -\\\hline

		\texttt{VELOCITY\_REQUEST\_HANDLER\_DISCARD\_PROB} & element from interval $[0,1]$ & Discard probability  for Velocity Request Handler to use & -\\\hline

		\texttt{ACCELERATION\_REQUEST\_HANDLER\_DISCARD\_PROB} & element from interval $[0,1]$ & Discard probability  for Acceleration Request Handler to use & -\\\hline
		
		
		\texttt{POSITION\_REQUEST\_HANDLER\_MODIFIER} & list of vector modifiers (see \ref{sec:vectorModifiers}) & List of vector modifiers for Position Request Handler to use & -\\\hline

		\texttt{VELOCITY\_REQUEST\_HANDLER\_MODIFIER} & list of vector modifiers (see \ref{sec:vectorModifiers}) & List of vector modifiers for Velocity Request Handler to use & -\\\hline

		\texttt{ACCELERATION\_REQUEST\_HANDLER\_MODIFIER} & list of vector modifiers (see \ref{sec:vectorModifiers}) & List of vector modifiers for Acceleration Request Handler to use & -\\\hline		
				
	\end{tabular}
	\caption{Variables in the main project file}\label{tab:mainvars2}
\end{sidewaystable}
\enlargethispage*{2cm}
\thispagestyle{empty}
\clearpage

\subsubsection{Activation Sequence Generators}

There are two Activation Sequence Generators (ASGs). A synchronous ASG and an asynchronous ASG.

To use the synchronous ASG one only needs to set the variable \texttt{ASG}=\texttt{SYNCHRONOUS}. No further variables need to be set.
To use the asynchronous ASG one needs to set \texttt{ASG}=\texttt{ASYNCHRONOUS}. Furthermore one needs to set the following variables:
\begin{itemize}
\item \texttt{ASYNC\_ASG\_SEED}: Seed for the ASG
\item \texttt{ASYNC\_ASG\_PART\_P}: The higher this is, the more robots are activated for each event. Goes from 0.0 to 1.0.
\item \texttt{ASYNC\_ASG\_TIME\_P}: The lower this is the smaller is the time difference between events. Be careful with very high values as buffer overflows might happen. Goes from 0.0 to 1.0.
\end{itemize}

\subsubsection{Vector Modifiers}\label{sec:vectorModifiers}

Vector modifiers are objects that change a given vector in a specific way. In addition to the to be modified vector, they may be given a so called \emph{reference vector}. How exactly this reference vector influences the modification of the given vector is due to the respective implementation.

Such vector modifiers are used by vector request handlers (e.g. position request handler). Requests to these handlers consist of a three dimensional vector (e.g. position the robot wants to be “beamed“ to). To model different aspects like inaccuracy or certain restrictions, vector request handlers may be given a list of vector modifiers. Before granting a request, every vector modifier in this list will be applied to the requested vector (e.g. the new position a robot requested) using the current value of the corresponding robot attribute (e.g. the robot's current position) as the reference vector.

The list of vector modifiers is a (not necessarily nonempty) list, i.\,e. 
\begin{center}\scriptsize
	\texttt{(VECTOR\_MODIFIER\_1);(VECTOR\_MODIFIER\_2);...}
\end{center}
The order of the elements of this list is important. If no Vector Modifier shall be used for the corresponding Request Handler, then use \texttt{VECTOR\_MODIFIERS=}''''.

An element \texttt{VECTOR\_MODIFIER\_k} of the Vector Modifier list is a tuple, defined as follows:
\begin{center}\scriptsize
	\texttt{VECTOR\_MODIFIER\_k=(VECTOR\_MODIFIER\_TYPE,VECTOR\_MODIFIER\_PARAM\_1,VECTOR\_MODIFIER\_PARAM\_2,..)}
\end{center}
The number and types of paramters like \texttt{VECTOR\_MODIFIER\_PARAM\_1,\newline VECTOR\_MODIFIER\_PARAM\_2,\dots} depends on the corresponding type of the Vector Modifier. Currently there are the following types of Vector Modifiers:
\begin{itemize}
	\item VectorDifferenceTrimmer
	\item VectorTrimmer
	\item VectorRandomizer
\end{itemize}
I.\,e. the value of \texttt{VECTOR\_MODIFIER\_TYPE} needs to be \texttt{VECTOR\_DIFFERENCE\_TRIMMER, VECTOR\_TRIMMER} or \texttt{VECTOR\_RANDOMIZER}.

\paragraph{VectorDifferenceTrimmmer} If \texttt{VECTOR\_MODIFIER\_TYPE} is equal to\newline \texttt{VECTOR\_DIFFERENCE\_TRIMMER}, then the following parameters are expected:
\begin{enumerate}
	\item \texttt{length} of type double: Vector request handlers that use vector difference trimmers will ensure that the robot's request does not differ too much from the current value of the robot's corresponding attribute. If the difference is larger than \texttt{length}, the robot's request is changed such that the difference's norm matches exactly \texttt{length} afterwards. For example, a position request handler using a vector difference trimmer with parameter $\texttt{length}=1$ will ensure that the robot can move only inside the unit sphere around it's current position. If the robot requested a position outside of this unit sphere, it would be moved on the intersection of the unit sphere's boundary and the line from the robot's current position to the requested position.
\end{enumerate}
I.\,e. an element of the \texttt{VECTOR\_MODIFIERS}-list of type \texttt{VECTOR\_DIFFERENCE\_TRIMMER} may look like: \texttt{(VECTOR\_DIFFERENCE\_TRIMMER,5.2)}.

\paragraph{VectorTrimmer} If vector modifier type equals \texttt{VECTOR\_TRIMMER}, then the following parameters are expected:
\begin{enumerate}
	\item \texttt{length} of type double: Every vector this vector modifier is applied to will be scaled such that it has length at most \texttt{length}. May be used for example by position request handlers to ensure that robots can not move out of the sphere of radius \texttt{length} around the origin of the global coordinate system.
\end{enumerate}
I.\,e. an element of the \texttt{VECTOR\_MODIFIERS}-list of type \texttt{VECTOR\_TRIMMER} may look like: \texttt{(VECTOR\_TRIMMER,10.0)}.

\paragraph{VectorRandomizer} If vector modifier type equals \texttt{VECTOR\_RANDOMIZER}, then the following parameters are expected:
\begin{enumerate}
	\item \texttt{seed} of type unsigned int: Seed used by randomization.
	\item \texttt{standard derivation} of type double: Derivation used by the vector randomizer. Applying this vector modifier to a vector will add a random vector distributed according to the multidimensional normal distribution with derivation \texttt{standard derivation} (i.e. every coordinate of the random vector is distributed according to N(0,\texttt{standard derivation})).
\end{enumerate}
I.\,e. an element of the \texttt{VECTOR\_MODIFIERS}-list of type \texttt{VECTOR\_DIFFERENCE\_TRIMMER} may look like: \texttt{(VECTOR\_DIFFERENCE\_TRIMMER,1,0.5)}.

\subsubsection{RobotControl}\label{sec:robotControl}
The \texttt{RobotControl} variable defines the class which should be used to control the robots (and in particular to control the views of the robots). Currently one of the following classes has be chosen:
\begin{enumerate}
	\item \texttt{UNIFORM\_ROBOT\_CONTROL}
	\item \texttt{ROBOT\_TYPE\_ROBOT\_CONTROL}
\end{enumerate}
Each class is explained in detail below. Note that each class expects certain class specific parameters.

\paragraph{UniformRobotControl}\label{sec:uniformRobotControl} This class assigns each robot the same view type. The concrete view type needs to be defined using a \texttt{VIEW} variable. The possible values for this variable (view types) are definied below (see \ref{sec:viewtypes}). E.\,g. you may to assign each robot global view to the world using \texttt{ROBOT\_TYPE\_ROBOT\_CONTROL="GLOBAL\_VIEW"}.

\paragraph{RobotTypeRobotControl}\label{sec:robotTypeRobotControl} This class assigns each robottype the same view type. Therefore robots with different robot types may have different view types. Currently there are two robot types:
\begin{enumerate}
	\item \texttt{MASTER}
	\item \texttt{SLAVE}
\end{enumerate}
To specify which view type should be used by each robot type, there must be variables of the form \textit{RobotType}\texttt{\_VIEW}. \\
The value of each variable has to be a view type (see \ref{sec:viewtypes}). Note that the view type parameters are also distinguished using the \textit{RobotType} prefix. E.\,g. you may specify 
\begin{center}
\texttt{MASTER\_VIEW="CHAIN\_VIEW"} \\
\texttt{MASTER\_CHAIN\_VIEW\_NUM\_ROBOTS="5"}
\end{center} to set the view for master robots to a chain view allowing the robots to see five neighbor robots. Note that exactly one view type should be defined for each robot type.

\subsubsection{ViewTypes}\label{sec:viewtypes}
The view type of a robot defines its vision model. Whenever a view type is expected you may use one of the following values:
\begin{enumerate}
	\item \texttt{GLOBAL\_VIEW}
	\item \texttt{COG\_VIEW}
	\item \texttt{CHAIN\_VIEW}
	\item \texttt{ONE\_POINT\_FORMATION\_VIEW}
	\item \texttt{SELF\_VIEW}
\end{enumerate}
Each view type is explained in detail below.

\paragraph{GLOBAL\_VIEW} Allows robots to see literally everything. There are no parameters expected.

\paragraph{COG\_VIEW} View model meant to be used for center of gravity algorithms, i.\,e. every robot can see every other robots position, velocity and acceleration. The coordinate-system and id of each robot is not visible. There are no parameters expected.

\paragraph{SELF\_VIEW} View model which allows robots to access every self-related information while disallowing to access any other information. There are no parameters expected.

\paragraph{CHAIN\_VIEW} View model meant to be used for robot chain related algorithms, i.\,e. every robot can see $k$ neighbor robots position. Besides this no more information is visible. When using this view type you have to specify the variable $k \in \mathbb{N}$ using the parameter variable \texttt{CHAIN\_VIEW\_NUM\_ROBOTS}.

\paragraph{ONE\_POINT\_FORMATION\_VIEW} View model meant to be used for one point formation algorithms, i.\,e. every robot can see every other robots position, velocity and acceleration only in a limited view radius $r$. The coordinate-system and id of each robot is not visible. When using this view type you have to specify the variable $r \in \mathbb{R}$ using the parameter variable \texttt{ONE\_POINT\_FORMATION\_VIEW\_RADIUS}.


\subsubsection{Example of a main project file}
A main project file may look like:
\lstset{language=tcl}
\begin{lstlisting}
# 
# Description about configuration.
#
	
	PROJECT_NAME="My Exciting Project"
	COMPASS_MODEL="NO_COMPASS"
	ROBOT_FILENAME="myrobots"
	OBSTACLE_FILENAME="myobstacle"
	STATISTICS_MODULE="0"
	ASG="ASYNCHRONOUS"
	ROBOT_CONTROL="ROBOT_TYPE_ROBOT_CONTROL"
	MASTER_VIEW="GLOBAL_VIEW"
	SLAVE_VIEW="ONE_POINT_FORMATION_VIEW"
	SLAVE_ONE_POINT_FORMATION_VIEW_RADIUS="5.0"
	
	CAMERA_POSITION="0,0,0"
	CAMERA_DIRECTION="1.5,0,0.5"
	
	MARKER_REQUEST_HANDLER_TYPE="STANDARD"
	STANDARD_MARKER_REQUEST_HANDLER_DISCARD_PROB="0.5"
	STANDARD_MARKER_REQUEST_HANDLER_SEED="1"

	TYPE_CHANGE_REQUEST_HANDLER_TYPE="NONE"
	# no additional variables needed

	POSITION_REQUEST_HANDLER_TYPE="VECTOR"
	VECTOR_POSITION_REQUEST_HANDLER_DISCARD_PROB="0.1"
	VECTOR_POSITION_REQUEST_HANDLER_SEED="3"
	VECTOR_POSITION_REQUEST_HANDLER_MODIFIER="(VECTOR_TRIMMER,1.5);(VECTOR_RANDOMIZER,5,2.5)"

	VELOCITY_REQUEST_HANDLER_TYPE="VECTOR"
	VECTOR_VELOCITY_REQUEST_HANDLER_DISCARD_PROB="0.1"
	VECTOR_VELOCITY_REQUEST_HANDLER_SEED="3"
	VECTOR_VELOCITY_REQUEST_HANDLER_MODIFIER="(VECTOR_TRIMMER,1.5);(VECTOR_RANDOMIZER,5,2.5)"
\end{lstlisting}


\subsection{Robot file}
The robotfile uses a csv-compatible format.
Therefore the information for one robot has to be saved in exactly one line of the file.
Each line contains the following data. The order of this data is important!

You can declare specific lines as comments by setting \# as the first sign of
the corresponding line.
\begin{itemize}
	\item ID-number
	\item initial position ($x,y,z$)
	\item initial type (for instance master, slave,$\ldots$)
	\item initial velocity ($x,y,z$)
	\item initial acceleration ($x,y,z$)
	\item initial status (maybe sleeping or ready; still has to be specified more precisely)
	\item initial marker information (still has to be specified)
	\item algorithm to use (shortcut for an algorithm; still needs to be specified)
	\item color (using this color a robot is marked for instance for a special treatment during the visualization; this color isn't used anywhere else); Integers correspond to the following colors: 0 green, 1 blue, 2 cyan, 3 red, 4 magenta, 5 yellow, 6 white, 7 black, 8 orange, 9 purple
	\item coordinate system axes (triple $x_1,x_2,x_3,y_1,y_2,y_3,z_1,z_2,z_3$; this field will be left empty, if axes are supposed to be generated uniformly at random)
\end{itemize}
The first line always is (column headers):
\begin{lstlisting}
	  "ID","x-position","y-position","z-position","type","x-velocity","y-velocity","z-velocity","x-acceleration","y-acceleration","z-acceleration","status","marker-info","algorithm","color","x-axis-1","x-axis-2","x-axis-3","y-axis-1","y-axis-2","y-axis-3","z-axis-1","z-axis-2","z-axis-3"
\end{lstlisting}
Each non-number is quoted and each number may be quoted.

\subsubsection{Example of a robot file}
\begin{lstlisting}
	"ID","x-position","y-position","z-position","type","x-velocity","y-velocity","z-velocity","x-acceleration","y-acceleration","z-acceleration","status","marker-info","algorithm","color","x-axis-1","x-axis-2","x-axis-3","y-axis-1","y-axis-2","y-axis-3","z-axis-1","z-axis-2","z-axis-3"
	0,5.3,9.2,6.4,"master",1.5,2.5,3.5,1.5,2.5,3.5,"sleeping",0,0,0,1,0,0,0,1,0,0,0,1
	1,"2.5","4.2","8.8","slave",1.5,2.5,3.5,1.5,2.5,3.5,"ready",0,1,0,1,0,0,0,1,0,0,0,1
\end{lstlisting}

\subsection{Obstacle file}
Like the robot file the obstacle file uses a csv-compatible format. 
Therefore the information for one robot has to be saved in exactly one line of the file.
Each line contains the following data. The order of this data is important!

You can declare specific lines as comments by setting \# as the first sign of
the corresponding line.
\begin{itemize}
	\item type (marker, sphere or box)
	\item position $(x,y,z)$
	\item marker information (still needs to be specified)
	\item $x/y/z$-lengths or radius (depending on type)
\end{itemize}

The first line always is (column headers):
\begin{lstlisting}
"type","x-position","y-position","z-position","marker-info","size-info","",""
\end{lstlisting}
Each non-number is quoted.


\subsubsection{Example of an obstacle file}
\begin{lstlisting}
"type","x-position","y-position","z-position","marker-info","size-info","",""
"box",2.0,3.0,4.0,0,1.0,2.0,3.0,
"sphere",3.4,5.2,5.1,0,5.0,"",""
"marker",3.5,1.4,5.1,0,"","",""
\end{lstlisting}
As you can already see in the example, if the type of an obstacle is sphere, then the last two values must be empty, i.\,e. '',''. Analoguos, if the type is marker, the last three values must be empty, i.\,e. '','',''.
